Network Fusion through CDRP Model\\ Decomposition and Assembly

### 2020-01-10 paper/paper.tex
some adjustments

To satisfy the above constraints, we follow the previous work of Wang \textit{et al.} \cite{wang2018interpret} to use critical data routing path (CDRP) for generating the input.
CDRP clustering allow us to decompose a large network model into several disjoint small ones.
To have overlapping classifiers, we decompose two different networks on the same dataset and fine tuning the CDRP models into binary classifiers.
We use image classification task on CIFAR-100 and VGG16 of type C and D for demonstration.

\paragraph{Remark.} We do not need to provide the subclasses in the input, because they can be derived from the initial set of classifiers.
Given encodings under the initial classifier set, samples share the same encoding can be considered as a subclass, and there is no need to further split those subclasses.
For example, in the intro example in Figure~\ref{fig:assemble-example-venn}, there is no need to distinguish $A_{11}$ and $A_{12}$ and we can group them as a single subclass.

\subsection{CDRP Decomposition and Clustering}
Guideline:
\begin{itemize}
	\item How to train gates, the functionality of the gates, how to interpret gate weight (lambda) feature of samples (similar to original paper)
	\item How to get decomposed classifiers: clustering
	\item Why clustered CDRP models might work (PCA visualization demonstration) (not performance analysis, inference experiments later)
	\item Why fine tuning the clustered CDRP model into binary classifiers, and how we fine tune them
\end{itemize}

Critical Data Routing Path is a method to find a path containing critical nodes, whose removal causes great preformance setback, from an existing trained model. We introduce CDRP to decomposite one existing pretrained model from selecting representing CDRPs in a given dataset.

\paragraph{CDRP extraction}

From the original model, we introduce a scalar control gate to every output channel of one layer, with nonnegative and sparse value. They will be multiplied to the layer’s output channelwise, resulting in the routing nodes. The nodes that are not suppressed, or have control gate weight greater zero, will be linked together to compose the routing paths. Then finding the critical data routing path can be reduced to optimize the weights of the control gates.

@@

\begin{algorithm}
	\caption{Distillation Guided Routing, introduced in paper \cite{wang2018interpret}}
	\label{alg:dgr}
	\begin{algorithmic}[1]
		\Require Input image $\boldsymbol{x}$, pretrained network $f_\theta(\cdot)$, control gates $\boldsymbol{\Lambda}$ initialized with 1, balanced parameter $\gamma$. Max iterations $T$, SGD optimizer
		\Ensure identified CDRPs representation $\boldsymbol{v}$
		\State original prediction class $i \gets \arg\max f_\theta(\boldsymbol{x})$
		\For{$t \gets 1$ to $T$}
			\State compute loss $cur\_loss$: $$\mathcal{L}\left(f_\theta(\boldsymbol{x}), f_\theta(\boldsymbol{x}; \boldsymbol{\Lambda})\right) + \gamma \sum_k |\boldsymbol{\lambda}_k|_1$$
			\State compute control gates gradients $\boldsymbol{\Lambda}$: $$\frac{\partial Loss}{\partial \boldsymbol{\Lambda}} = \frac{\partial \mathcal{L}}{\partial \boldsymbol{\Lambda}} + \gamma \times \mathrm{sign}(\boldsymbol{\Lambda})$$
			\State update $\boldsymbol{\Lambda}$ by SGD optimzier and clip $\boldsymbol{\Lambda}$ to be nonnegative
			\State new prediction class $j \gets \arg\max f_\theta(\boldsymbol{x};\boldsymbol{\Lambda})$
			\If{$i=j$} \Comment keep the prediction same
				\If{$cur\_loss$ is minimun}
					\State $\boldsymbol{v} \gets \mathrm{concatenate}(\boldsymbol{\Lambda})$
				\EndIf
			\EndIf
		\EndFor		
	\end{algorithmic}
\end{algorithm}

\paragraph{CDRP clustering}

@@

\begin{figure}[!htp]
	\centering
	\subfigure[Result of VGG-16(C)] {
		\includegraphics[width=\linewidth]{fig/vgg16c-cluster.png}
        \label{fig:vgg16c-cluster}
	}
	\subfigure[Result of VGG-19(D)] {
		\includegraphics[width=\linewidth]{fig/vgg16d-cluster.png}
        \label{fig:vgg16d-cluster}
    }
    \caption{Visualization of clustering results. The axes are the first two components of of PCA results on all extracted gate vectors. The figure in the top is the result of VGG-16 (type C) trained on cifar-100 dataset, and the one below is the result of VGG-16 (type D) fine-tuned on cifar-100 initialized with pretrained weights on ImageNet. Different colors represent different class labels in cifar-100 dataset, and the cluster centers are denoted as blue squares.}
    \label{fig:vgg-cluster-pca}
\end{figure}

@@

\paragraph{Model pruning and fine tuning}

@@

\section{Experiments}
\label{sec:experiments}
\subsection{Verifying CDRP Models of VGGNets trained on CIFAR-100}
In this section, we perform K-Means on the CDRP control gates of each sample. For each cluster, we get a binary-classifier to determine whether the test sample should be in the cluster. Then, we fine tune each classifier in the CIFAR-100 dataset.

\paragraph{Clusters.} For VGGNet, we generate $1.5$ clusters per class.
Then, we get $14$ clusters from $5$ classes, among which $7$ clusters are from VGG16-C and $7$ clusters from VGG16-D.

The sample distribution of each cluster is shown in Figure~\ref{fig:vgg-cluster}.

\begin{figure}[!htp]
    \centering
\subfigure[Clusters from VGG-C]{
    \begin{minipage}[t]{1\linewidth}
        \includegraphics[height=1.5in,width=2.8in]{fig/chart1.pdf}
        \label{fig:net2a}
    \end{minipage}
}

    %\vfill
\subfigure[Clusters from VGG-D]{
    \begin{minipage}[t]{1\linewidth}
        \includegraphics[height=1.5in,width=2.8in]{fig/chart2.pdf}
        \label{fig:net2b}
    \end{minipage}
}
    \caption{Sample Distribution of Clusters from Different Types of VGGNets}
    \label{fig:vgg-cluster}

@@

\subsection{Fusing VGGNets for Specific Tasks through CDRP Assembly}
In this section, we implement the heuristic search in \cref{tab:heuristic-search} to fuse the classifiers. According to the 0/1 result of each classifier, we can get an unique encoding for each sample, which corresponds to a unqiue label so that we can determine the class of each test sample.

\paragraph{Fusion}We implement \cref{tab:heuristic-search} to fuse the classifiers since we want to finish a specific task of classifying among five classes, we set a upper bound of the number of classifiers $k=4$. 
After we select these classifiers with this \cref{tab:heuristic-search}, different results of each classifier ($0$ and $1$ represent the sample is or isn't in the cluster respectively) correspond to a unique label.
In our implementation, the algorithm choose four classifiers: 3,4,1,2 in order. 
The accuracy of the four classifiers with different prune ratio is shown in \cref{tab:classifier-accuracy}.The test set has also been balanced to make the number of positive and negative samples nearly $1:1$.

>>>>

To satisfy the above constraints, we follow the previous work of Wang \textit{et al.} \cite{wang2018interpret} to use critical data routing path (CDRP) for generating the input.
CDRP clustering allow us to decompose a large network model into several disjoint small ones.
To have overlapping classifiers, we decompose two different networks on the same dataset and fine tuning the CDRP models into binary classifiers.
We use image classification task on CIFAR-100 and VGG16 of type C and type D for demonstration.

\paragraph{Remark.} We do not need to provide the subclasses in the input, because they can be derived from the initial set of classifiers.
Given encodings under the initial classifier set, samples share the same encoding can be considered as a subclass, and there is no need to further split those subclasses.
For example, in the intro example in \Cref{fig:assemble-example-venn}, there is no need to distinguish $A_{11}$ and $A_{12}$ and we can group them as a single subclass.

\subsection{CDRP Decomposition and Clustering}
Critical Data Routing Path is a method to find a path containing critical nodes, whose removal causes great preformance setback, from an existing trained model. We introduce CDRP to decomposite one existing pretrained model from selecting representing CDRPs in a given dataset.

\paragraph{CDRP extraction.}

From the original model, we introduce a scalar control gate to every output channel of one layer, with nonnegative and sparse value. They will be multiplied to the layer's outputs channelwise, resulting in the routing nodes. The nodes that are not suppressed, or have control gate weight greater zero, will be linked together to compose the routing paths. Then finding the critical data routing path can be reduced to optimize the weights of the control gates.

@@

\begin{algorithm}
	\caption{Distillation Guided Routing, from \cite{wang2018interpret}}
	\label{alg:dgr}
	\begin{algorithmic}[1]
		\Require Input image $\boldsymbol{x}$, pretrained network $f_\theta(\cdot)$, control gates $\boldsymbol{\Lambda}$ initialized with 1, balanced parameter $\gamma$. Max iterations $T$, SGD optimizer
		\Ensure identified CDRPs representation $\boldsymbol{v}$
		\State original prediction class $i \gets \arg\max f_\theta(\boldsymbol{x})$
		\For{$t \gets 1$ to $T$}
			\State compute loss $cur\_loss$: $$\mathcal{L}\left(f_\theta(\boldsymbol{x}), f_\theta(\boldsymbol{x}; \boldsymbol{\Lambda})\right) + \gamma \sum_k |\boldsymbol{\lambda}_k|_1$$
			\State compute control gates gradients $\boldsymbol{\Lambda}$: $$\frac{\partial Loss}{\partial \boldsymbol{\Lambda}} = \frac{\partial \mathcal{L}}{\partial \boldsymbol{\Lambda}} + \gamma \times \mathrm{sign}(\boldsymbol{\Lambda})$$
			\State update $\boldsymbol{\Lambda}$ by SGD optimzier and clip $\boldsymbol{\Lambda}$ to be nonnegative
			\State new prediction class $j \gets \arg\max f_\theta(\boldsymbol{x};\boldsymbol{\Lambda})$
			\If{$i=j$} \Comment keep the prediction same
				\If{$cur\_loss$ is minimun}
					\State $\boldsymbol{v} \gets \mathrm{concatenate}(\boldsymbol{\Lambda})$
				\EndIf
			\EndIf
		\EndFor		
	\end{algorithmic}
\end{algorithm}

\paragraph{CDRP clustering.}

@@

\begin{figure}
	\centering
	\subfigure[Result of VGG-16(C)] {
		\includegraphics[width=\linewidth]{fig/vgg16c-cluster.png}
        \label{fig:vgg16c-cluster}
	}
	\subfigure[Result of VGG-19(D)] {
		\includegraphics[width=\linewidth]{fig/vgg16d-cluster.png}
        \label{fig:vgg16d-cluster}
    }
    \caption{Visualization of clustering results. The axes are the first two components of of PCA results on all extracted gate vectors. The figure in the top is the result of VGG-16 (type C) trained on cifar-100 dataset, and the one below is the result of VGG-16 (type D) fine-tuned on cifar-100 initialized with pretrained weights on ImageNet. Different colors represent different class labels in cifar-100 dataset, and the cluster centers are denoted as blue squares.}
    \label{fig:vgg-cluster-pca}
\end{figure}

@@

\paragraph{Model pruning and fine tuning.}

@@

\section{Experiments}
In this section, we implement the CDRP decomposition and assembly framework instantiated in \cref{sec:cdrp}.
We perform K-Means on the CDRP control gates of each sample and extract binary classifiers from clusters' corresponding CDRP model.
We apply the search \cref{tab:heuristic-search} in \cref{sec:assembly-algorithm} to fuse those classifiers, and we test the performance of each binary classifiers and the combined model.

\label{sec:experiments}
\subsection{Verifying CDRP Models of VGGNets trained on CIFAR-100}
\paragraph{Clusters.} For VGGNet, we generate $1.5$ clusters per class.
Then, we get $14$ clusters from $5$ classes, among which $7$ clusters are from VGG16-C and $7$ clusters from VGG16-D.

The sample distribution of each cluster is shown in \Cref{fig:vgg-cluster}.

\begin{figure}[h]
    \centering
    \begin{minipage}{0.99\linewidth}
		\subfigure[Clusters from VGG-C]{
			\includegraphics[width=\linewidth]{fig/chart1.pdf}
			\label{fig:net2a}
		}
    \end{minipage}

    %\vfill
    \begin{minipage}{0.99\linewidth}
		\subfigure[Clusters from VGG-D]{
			\includegraphics[width=\linewidth]{fig/chart2.pdf}
			\label{fig:net2b}
		}
    \end{minipage}
    \caption{Sample Distribution of Clusters from Different Types of VGGNets}
    \label{fig:vgg-cluster}

@@

\subsection{Fusing VGGNets for Specific Tasks through CDRP Assembly}

\paragraph{Fusion.} We implement \cref{tab:heuristic-search} to fuse the classifiers since we want to finish a specific task of classifying among five classes, we set a upper bound of the number of classifiers $k=4$. 
After we select these classifiers with this \cref{tab:heuristic-search}, different results of each classifier ($0$ and $1$ represent the sample is or isn't in the cluster respectively) correspond to a unique label.
In our implementation, the algorithm choose four classifiers: 3,4,1,2 in order. 
The accuracy of the four classifiers with different prune ratio is shown in \cref{tab:classifier-accuracy}.The test set has also been balanced to make the number of positive and negative samples nearly $1:1$.
